---
title: "Data 607 - Homework 1"
author: "Peter Kowalchuk"
date: "August 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Load livrary to scrape data from University of California Irvine website

```{r}
library(curl)
```

Load data original data from website into variable originalDataCSV

```{r}
data_url <- 'https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data'
original <- curl(data_url)
originalDataCSV<-readLines(original)
```


Store original CSV data into an R dtaframe
```{r}
originalData<-read.csv(text=originalDataCSV,header=FALSE)
```


See first few rows of data set
```{r}
head(originalData)
```

Create array with names for all atributes/columns
```{r}
atributes<-c('classes','cap-shape','capsurface','capcolor','bruises','odor','gillattachment','gillspacing','gillsize','gillcolor','stalkshape','stalkroot','stalksurfaceabovering','stalksurfacebelowring','stalkcolorabovering','stalkcolorbelowring','veiltype','veilcolor','ringnumber','ringtype','sporeprintcolor','population','habitat')
```

Assign Ccolumn names to data set
```{r}
colnames(originalData)<-atributes
head(originalData)
```

Subset the data frame to include the classes column (edible or not), and four other columns. Selected were: odor, gill-color, stalk-root and habitat
```{r}
selectedData<-originalData[c('classes','odor','gillsize','stalkshape','habitat')]
head(selectedData)
```

Next we replace the abbreviations used in the dataset with meaningful names.

```{r}

classes<-data.frame(sapply(selectedData$classes,function(x) { gsub("e","edible",x)}))
classes<-data.frame(sapply(classes,function(x) { gsub("p","poisonous",x)}))

odor<-data.frame(sapply(selectedData$odor,function(x) { gsub("a","almond",x)}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="l","anise",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="c","creosote",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="y","fishy",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="f","foul",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="m","musty",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="n","none",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="p","pungent",as.character(x))}))
odor<-data.frame(sapply(odor,function(x) {ifelse(x=="s","spicy",as.character(x))}))

gillsize<-data.frame(sapply(selectedData$gillsize,function(x) {ifelse(x=="b","broad",as.character(x))}))
gillsize<-data.frame(sapply(gillsize,function(x) {ifelse(x=="n","narrow",as.character(x))}))

stalkshape<-data.frame(sapply(selectedData$stalkshape,function(x) {ifelse(x=="e","enlarging",as.character(x))}))
stalkshape<-data.frame(sapply(stalkshape,function(x) {ifelse(x=="t","tapering",as.character(x))}))

habitat<-data.frame(sapply(selectedData$habitat,function(x) {ifelse(x=="g","grasses",as.character(x))}))
habitat<-data.frame(sapply(habitat,function(x) {ifelse(x=="l","leaves",as.character(x))}))
habitat<-data.frame(sapply(habitat,function(x) {ifelse(x=="m","meadows",as.character(x))}))
habitat<-data.frame(sapply(habitat,function(x) {ifelse(x=="p","paths",as.character(x))}))
habitat<-data.frame(sapply(habitat,function(x) {ifelse(x=="u","urban",as.character(x))}))
habitat<-data.frame(sapply(habitat,function(x) {ifelse(x=="w","waste",as.character(x))}))
habitat<-data.frame(sapply(habitat,function(x) {ifelse(x=="d","woods",as.character(x))}))

finalData<-data.frame(classes=classes,odor=odor, gillsize=gillsize, stalkshape=stalkshape,habitat=habitat)
colnames(finalData)<-c('classes','odor','gillsize','stalkshape','habitat')
head(finalData)
tail(finalData)

```



